---
title: "Analysis of Binary Classification Results Using Imputed Drug Signatures"
author: "Rachel Hodos"
date: '`r substr(date(), 5, 10)`, 2017'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = '../../')
```

```{r setup2, echo=FALSE, message=FALSE, warning=FALSE}

library(ggplot2)
library(reshape2)
library(plyr)
library(RColorBrewer)
library(ggpubr)

options(error=recover)
library(RSQLite)
source('R/src/DataProc.R')
source('R/src/Utils.R')

SplitOutcome = function(df){
  df$category = sapply(as.character(df$outcome), function(x) unlist(strsplit(x, split='[.]'))[[1]])
  df$outcome = sapply(as.character(df$outcome), function(x) unlist(strsplit(x, split='[.]'))[[2]])
  return(df)
}

FilterByAUC = function(df, threshold=0.5){
  return(subset(df, AUC.full > threshold | AUC.obs > threshold))
}
```

Load the results
```{r load_data}
load(ResultsDir('ATC_and_target_prediction/results_ROC_counts_params.RData'))
```

Do some processing...
```{r process_Rmeas}

# Melt AUC results
R = melt(ROC)
names(R) = c('AUC', 'eval', 'model', 'subset', 'feature', 'outcome')

# Split outcome into outcome and category
R = SplitOutcome(R)

# Subset to top three represented ATCs
R = subset(R, outcome %in% c('L','C','D') | category == 'Target')

# Subset to evaluations on measured signatures and reformat data
Rmeas = RemoveDfColumns(subset(R, eval == 'eval_meas'), 'eval')
Rmeas$AUC[is.na(Rmeas$AUC)] = 0.5
Rmeas = dcast(Rmeas, model + feature + outcome + category ~ subset, value.var='AUC')
Rmeas = suppressWarnings(ChangeColumnName(Rmeas, from=c('full','obs'), to=c('AUC.full', 'AUC.obs')))
Rmeas$diff = Rmeas$AUC.full - Rmeas$AUC.obs
idx = Rmeas$category == 'ATC'
Rmeas$outcome[idx] = paste0('ATC ' , Rmeas$outcome[idx])

# Results where AUCs were both below this in the two comparisons are thrown out
threshold = 0.5 
Rmeas = FilterByAUC(Rmeas, threshold)
```

#### Analysis of overall results
Look at the overall difference between the two groups:
```{r overall_stats}
t.test(x=Rmeas$AUC.full, y=Rmeas$AUC.obs, paired=TRUE)
```

Then count the number of cases where there is a positive change in AUC:
```{r overall_counts}
nPos = length(which(Rmeas$diff > 0)) 
nTot = nrow(Rmeas)
print(sprintf('%d out of %d increased AUC', nPos, nTot))
```

#### Analysis of results per model
Plot distributions.
```{r deltas_per_model}
print(ggplot(Rmeas, aes(x=model, y=diff, group=model, fill=model)) +
        geom_boxplot() + ggtitle('Improvement in AUC per model') + stat_compare_means())
```

Check significance of deltas per model.
```{r significance_per_model}
print(lapply(split(Rmeas, Rmeas$model), function(x) t.test(x$AUC.full, x$AUC.obs, paired=TRUE)))
```

#### Analysis of results per signature type
```{r plot_deltas_per_sig}
numSigs = c(MCF7=1505, VCAP=1368, PC3=1340, A375=1168, A549=1139,
            HA1E=1127, HT29=1022, HCC515=934, HEPG2=798, NPC=441)
Rmeas$num_sigs = numSigs[Rmeas$feature]

ggplot(Rmeas, aes(x=reorder(feature, -diff, FUN=median), y=diff, group=feature, fill=num_sigs)) +
  geom_hline(yintercept = 0, color='grey', lwd=1) +
  geom_boxplot(alpha=0.9) +
  ggtitle(sprintf('Deltas per feature, threshold = %0.1f', threshold)) +
  theme_bw() + ylim(c(-0.5,0.5)) + scale_fill_gradientn(colours=brewer.pal(5,'YlOrRd')) +
  theme(text = element_text(size=22), axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=24, hjust=0.5),
        legend.justification='bottom', legend.position='bottom') +
  labs(fill='# Signatures  \n  Measured\n', x='', y=expression(AUC [impute] - AUC [orig]),
       title='Improvement in AUC per\n Drug Signature Type')
ggsave(PlotDir('DeltaAUC_per_Feature.svg'), height=7.7, width=8)
```
```{r signature_stats}
p_feature = sapply(split(Rmeas, Rmeas$feature), function(x) t.test(x$AUC.full, x$AUC.obs, paired=TRUE)$p.value)
print(adjp_feature <- sort(p.adjust(p_feature, method='BH')))
```

```{r plot_deltas_per_outcome}
ggplot(Rmeas, aes(x=reorder(outcome, -diff, FUN=median), y=diff, group=outcome, fill=category)) + ylim(c(-0.5,0.5)) +
  geom_hline(yintercept = 0, color='grey', lwd=1) + geom_boxplot(alpha=0.8) + scale_fill_manual(values=c('#CC9966','#009999')) +
  ggtitle(sprintf('Deltas per outcome, threshold = %0.1f', threshold)) + theme_bw() +
  labs(x='', y=expression(AUC [impute] - AUC [orig]), title='Improvement in AUC per\n Prediction Task',
       fill='Category') +
  theme(text = element_text(size=22), axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=24, hjust=0.5),
        legend.justification='bottom', legend.position='bottom')
ggsave(PlotDir('DeltaAUC_per_Outcome.svg'), height=7, width=8)
```

```{r outcome_stats}
p_outcome = lapply(split(Rmeas, Rmeas$outcome), function(x) t.test(x$AUC.full, x$AUC.obs, paired=TRUE)$p.value)
print(adjp_outcome <- sort(p.adjust(p_outcome, method='BH')))

ddply(Rmeas, 'outcome', summarise, median.diff= median(diff))
```

```{r process_q2_data}
load(ResultsDir('ATC_and_target_prediction/new_ROC_and_counts_with_correct_AUC_imp_values.RData'))

Outcome2Category = function(outcome){
  return(sapply(as.character(outcome), function(x) unlist(strsplit(x, split='[.]'))[[1]]))
}

# Reformat results
R = melt(ROC)
R = dcast(R, L1 + L2 + L3 + L4 ~ L5)
names(R) = c('outcome','feature','subset','model', 'AUC_imp', 'AUC_meas')
R$category = Outcome2Category(R$outcome)
R$outcome = sapply(R$outcome, function(x) unlist(strsplit(x, split='[.]'))[[2]])

# Merge R with counts
C = melt(counts)[,-3]
names(C) = c('count','variable','feature','outcome')
C = dcast(C, feature + outcome  ~ variable, value.var='count')
C$outcome = sapply(C$outcome, function(x) unlist(strsplit(x, split='[.]'))[[2]])
RC = merge(R, C, all=TRUE, by=c('outcome','feature'))
RC = subset(RC, category=='Target' | outcome %in% c('L','C','D')) 
print(sprintf('Starting with %d experiments', nrow(RC)))

# Filter by two different thresholds to avoid 'non-signal' cases
RC = subset(RC, AUC_imp > 0.5 | AUC_meas > 0.5 )
print(sprintf('After filtering by AUC > 0.5, %d experiments remaining', nrow(RC)))

RC = subset(RC, nPos_imp >= 3 & nPos_meas >= 3)
print(sprintf('After filtering by num labels >= 3, %d experiments remaining', nrow(RC)))

Targets = subset(RC, category == 'Target')
ATC = subset(RC, category == 'ATC')
```

Run stats for ATC code prediction:
```{r ATC_stats}
t.test(ATC$AUC_imp, ATC$AUC_meas, paired=TRUE)
```

Run similar stats for target prediction:
```{r target_stats}
t.test(Targets$AUC_imp, Targets$AUC_meas, paired=TRUE)
```

```{r HT29}
H = subset(Targets, feature == 'HT29')
print(sprintf('mean improvement by HT29 imputed signatures: %0.2f',
              mean(H$AUC_imp - H$AUC_meas)))
```


Compute median value across three models to simplify plots:
```{r median_per_model}
print(sprintf('Size of RC before: %d', nrow(RC)))
RC = ddply(RC, c('outcome','feature','category','subset'),
          summarize, AUC_imp=median(AUC_imp), AUC_meas=median(AUC_meas))
print(sprintf('Size of RC after: %d', nrow(RC)))
Targets = subset(RC, category == 'Target')
ATC = subset(RC, category == 'ATC')
print(sprintf('Size of ATC and Targets: %d, %d', nrow(ATC), nrow(Targets)))
```

Make ATC scatter (Figure 5D):
```{r ATC_scatter}
ATC$outcome = revalue(ATC$outcome, c('C'='C: Cardiovascular', 'D'='D: Dermatological',
                                     'L'='L: Antineoplastic/Immunomodulatory'))
ggplot(ATC, aes(x=AUC_meas, y=AUC_imp, color=outcome, label=feature)) + 
  geom_abline(slope=1, intercept=0, color='DarkGrey', lty='dashed') + 
  geom_point(size=7, alpha=0.6) +
  xlim(c(0,1)) + ylim(c(0,1)) +
  geom_text(size=5, color='black', check_overlap=TRUE) + theme_bw() + 
  guides(color=guide_legend(title='ATC Code', nrow=3)) + 
  xlab('AUC on measured signatures') + ylab('AUC on predicted signatures') + 
  theme(axis.text=element_text(size=18), axis.title=element_text(size=22),
        legend.text=element_text(size=18), legend.title=element_text(size=22),
        legend.position='bottom', legend.justification='bottom')
ggsave(PlotDir('ATC_code_imputed_vs_measured.svg'), width=8, height=7)
```

Make Target scatter (Figure 5E):
```{r Target_scatter}
ggplot(Targets, aes(x=AUC_meas, y=AUC_imp, color=outcome, label=feature)) + 
  geom_abline(slope=1, intercept=0, color='DarkGrey', lty='dashed') + geom_point(size=7, alpha=0.6) +
  xlim(c(0,1)) + ylim(c(0,1)) +
  geom_text(size=5, color='black', check_overlap=TRUE) + theme_bw() + 
  guides(color=guide_legend(title='Target', nrow=2)) +
  xlab('AUC on measured signatures') + ylab('AUC on predicted signatures') + 
    theme(axis.text=element_text(size=18), axis.title=element_text(size=22),
        legend.text=element_text(size=18), legend.title=element_text(size=22),
        legend.position='bottom', legend.justification='bottom')
ggsave(PlotDir('Target_imputed_vs_measured.svg'), width=8, height=7)
```

